---
title: "Reproducible Research Project One"
author: "Amanda Mae"
date: "7/1/2019"
output: html_document
---
#This is the code for Reproducible Research, Course Project One

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Step One: Code for reading in the dataset and/or processing the data

```{r}
setwd("~/Desktop/R Respository/Reproducible Research")
activity <- read.csv("activity.csv", header = TRUE, sep = ",")
```

###Check that the file was read in correctly

```{r}
dim(activity)
variable.names(activity)
str(activity)
```

##Step Two: Histogram of the total number of steps taken each day

*Step TwoA: Create variable steps per day*

```{r}
steps_day<-with(activity,tapply(steps,date,sum,na.rm=TRUE))
```

*Step TwoB: Create and Save Histogram*

```{r echo=FALSE}
hist(steps_day,col = "blue",xlab = "Total Steps",ylab = "Days",main = "Frequency of Total Number of Steps per Day")
```



##Step Three: Mean and median number of steps taken each day

```{r}
mean(steps_day)
median(steps_day)
```

*OR (produces both)*


```{r}
summary(steps_day)
```


#Step Four: Time series plot of the average number of steps taken

```{r, echo=FALSE}
avg_steps<-with(activity,tapply(steps,interval,mean,na.rm=TRUE))
intervals<-unique(activity$interval)
Time_Series_Data<-data.frame(cbind(avg_steps,intervals))
plot(Time_Series_Data$intervals,Time_Series_Data$avg_steps,type = "l",xlab = "Intervals", 
     ylab = "Average Steps",main = "Average Steps per Interval")
```

#Step Five: The 5-minute interval that, on average, contains the maximum number of steps

```{r}
index<-which.max(Time_Series_Data$avg_steps)
max<-Time_Series_Data[index,2]
max
```


#Step Six: Code to describe and show a strategy for imputing missing data

*How many missing values do we have?*
```{r}
sum(is.na(activity$steps))
```
*Replace missing values with average daily value*
```{r}
index<-which(is.na(activity$steps))
l<-length(index)
   steps_avg<-with(activity,tapply(steps,date,mean,na.rm=TRUE))
   na<-mean(steps_avg,na.rm = TRUE)
    for (i in 1:l) {
    activity[index[i],1]<-na
    }
```

*Check All Missing Values Have Been Replaced*
```{r}
sum(is.na(activity$steps))
str(activity)
```


#Step Seven: Histogram of the total number of steps taken each day after missing values are imputed

```{r}
avg_steps2<-with(activity,tapply(steps,date,sum,na.rm=TRUE))
hist(avg_steps2,col = "yellow",xlab = "Total Steps",ylab = "Days",main = "Frequency of Total Number of Steps per Day")
```


#Step Eight: Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
library(dplyr)
```

*Part One is Identifying and Tagging Weekends/Weekdays*

```{r}
activity$day <- weekdays(as.Date(activity$date))
head(activity)
activity$day_type <- ifelse(activity$day == "Saturday" | activity$day == "Sunday", "weekend", "weekday")
head(activity)
str(activity)
```

*Now Plot!*
```{r}
StepsPerInterval.weekend <- tapply(activity[activity$day_type == "weekend" ,]$steps, activity[activity$day_type == "weekend" ,]$interval, mean, na.rm = TRUE)

StepsPerInterval.weekday <- tapply(activity[activity$day_type == "weekday" ,]$steps, activity[activity$day_type == "weekday" ,]$interval, mean, na.rm = TRUE)

par(mfrow=c(1,2))

plot(as.numeric(names(StepsPerInterval.weekday)), 
     StepsPerInterval.weekday, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekdays)", 
     type = "l")

plot(as.numeric(names(StepsPerInterval.weekend)), 
     StepsPerInterval.weekend, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekends)", 
     type = "l")
```

#Step Nine: All Code in R Markdown!